---
title: "Stat 341 -- Homework 06"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(mosaic)      # this loads ggformula (for plotting), etc. too
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable() for tables
library(patchwork)   # for combining plots

# data wrangling
library(tidyverse)

# several packages for bayesian stuff -- more may come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)
library(CalvinBayes)



# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```


<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

```{r}
data(foxes)
foxes <- na.omit(foxes)
glimpse(foxes)
```

Interested in the affect of avgfood, groupsize on weight, avgfood on groupsize, and area on avgfood.

```{r}
foxes <- foxes |>
  mutate(avgfood_scaled = as.numeric(scale(avgfood))) |>
  mutate(groupsize_scaled = as.numeric(scale(groupsize))) |>
  mutate(area_scaled = as.numeric(scale(area))) |>
  mutate(weight_scaled = as.numeric(scale(weight)))

glimpse(foxes)
```

```{r}
model_descrip <- alist(
  # note variable name has to match actual data variable name
  weight_scaled ~ dnorm(mu, sigma),
  mu ~ beta0 + beta1 * avgfood_scaled + beta2 * groupsize_scaled + beta3 * area_scaled,
  beta0 ~ dnorm(mean = 0, sd = 1),
  beta1 ~ dnorm(mean = 0, sd = 2),
  beta2 ~ dnorm(mean = 0, sd = 2),
  beta3 ~ dnorm(mean = 0, sd = 2),
  sigma ~ dnorm(mean = 0, sd = 1)
)
```

```{r}
model_descrip2 <- alist(
  # note variable name has to match actual data variable name
  weight_scaled ~ dnorm(mu, sigma),
  mu ~ beta0 + beta1 * avgfood_scaled + beta2 * groupsize_scaled,
  beta0 ~ dnorm(mean = 0, sd = 1),
  beta1 ~ dnorm(mean = 0, sd = 2),
  beta2 ~ dnorm(mean = 0, sd = 2),
  sigma ~ dnorm(mean = 0, sd = 1)
)
```

```{r}
model_descrip3 <- alist(
  # note variable name has to match actual data variable name
  weight_scaled ~ dnorm(mu, sigma),
  mu ~ beta0 + beta1 * avgfood_scaled,
  beta0 ~ dnorm(mean = 0, sd = 1),
  beta1 ~ dnorm(mean = 0, sd = 2),
  sigma ~ dnorm(mean = 0, sd = 1)
)
```


```{r}
quap_fox_model <- quap(flist = model_descrip,
                          data = foxes)
```


```{r}
quap_fox_model2 <- quap(flist = model_descrip2,
                          data = foxes)
```

```{r}
quap_fox_model3 <- quap(flist = model_descrip3,
                          data = foxes)
```

```{r}
compare(quap_fox_model, quap_fox_model2 ,quap_fox_model3, func=WAIC )
```

The WAIC are also pretty close to one another but the initial normal model with 3 different parameters has the lowest score and thus has the lowest penalty in results. I had a third model with a WAIC score of 333.9 but I wasn't able to knit the file so I am still trouble shootinng. It may be it this new knit file.


