---
title: "Stat 341 -- Homework 03"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used


# data wrangling


# several packages for bayesian stuff -- more may come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)

# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```

<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## N1

```{r}
library(dagitty)
library(CalvinBayes)

# create the DAG object
causal_diagram_1 <- dagitty("dag{
  area -> avgfood -> weight;
  avgfood -> groupsize -> weight
}")
# plot it
gg_dag(causal_diagram_1)
```

## N2
A. 
avgfood is a mediator variable between area and weight.
groupsize is a mediator variable between avgfood and weight.  

B.
I would include both avgfood and groupsize variables since I don't think either are collider variables so I think they are helpful.

## N3

A.
How does age and sex of a countries' current population project their future population?

B.
age and sex are the predicting variables for population, the response variable.

C.
current population, number of births, number of deaths, number of immigrants, and number of emigrants.

D.
```{r}
# create the DAG object
causal_diagram_3 <- dagitty("dag{
  age -> number.of.births.minus.number.of.deaths.plus.number.of.immigrants.minus.number.of.emigrants -> population;
  sex ->  number.of.births.minus.number.of.deaths.plus.number.of.immigrants.minus.number.of.emigrants -> population
}")
# plot it
gg_dag(causal_diagram_3)
```


E.
the calculations for population based on births, deaths, immigrants, and emigrants is a mediator variable

F.
Seems like the researchers model was based around either age or sex and then modified by their population calculation and then modeled so the causal diagram should be consistent.

## N4

A. After fitting a model in homework 2 to look at if a movie is animated or not, I would think that if the genre is or isn't horror and if the movie was for kids would both be variables relevant to the proportion I was estimating. I think more kids movies and movies that aren't horror movies would be relevant to if the movie was animated or not.

B.

```{r}
# create the DAG object
causal_diagram_2 <- dagitty("dag{
  animation -> horror -> for_kids -> animated;
  horror -> animated
}")
# plot it
gg_dag(causal_diagram_2)
```


C.
for_kids is a mediator variable and horror is a confounder variable.

D.
I think both for_kids and horror should have been included in the model fitted in homework 02.
