---
title: "Stat 341 -- Homework 8"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(mosaic)      # this loads ggformula (for plotting), etc. too
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable() for tables
library(patchwork)   # for combining plots

# data wrangling
library(tidyverse)

# several packages for bayesian stuff -- more may come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)
library(CalvinBayes)


# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```


<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## 9E4
Explain the difference between the effective number of samples, n_eff as calculated by stan, and the actual number of samples.

"The effective number of samples is an estimate of the number of independent samples from the posterior distribution, in terms of estimating
some function like the posterior mean. (287)" n_eff is the length of a markov chain with no autocorrelationn whereas the actual number of camples are typically autocorrelated.

## 9E5
Which value should Rhat approach, when a chain is sampling the posterior distribution correctly?

Rhat should approach 1 when the chain is sampling the posterior distribution correctly.

## 9E6
Second PDF with drawing.

## 9E7
Second PDF with drawing.

## 9M3

```{r}
bugs <- read_csv('https://sldr.netlify.app/data/house_bugs.csv') |>
  na.omit() |>
  mutate(arthropod_div = arthropod.div) |>
  mutate(income_avg_z = income.avg.z) |>
  mutate(sqft_z = sqft.z) |>
  mutate(total_value_z = total.value.z)
```

```{r}
stan_data <- compose_data(bugs)

stan_program <- '
data {
// number of observations
int<lower=1> n;
// response
vector[n] arthropod_div;
// predictor
vector[n] income_avg_z;
// predictor
vector[n] sqft_z;
// predictor
vector[n] total_value_z;
}
parameters {
// std of response, single continuous value
real<lower=0> sigma;
real a;
real b;
real c;
real d;
}
model {
// vector of n values: expected arthropod.div for each observation
vector[n] mu;
// loop over the n cases in the dataset to estimate mu_i values
for (i in 1:n) {
mu[i] = a + (b*income_avg_z[i]) + (c*total_value_z[i]) + (d*sqft_z[i]);
}
// prior for both intercepts
a ~ normal(20, 10);
// prior for both slopes
b ~ normal(25, 5);
c ~ normal(25, 5);
d ~ normal(10, 10);
// prior for sigma
sigma ~ normal(20, 5);
// defining likelihood in terms of mus and sigma
arthropod_div ~ normal(mu, sigma);
}
'
```

```{r, message=FALSE, results='hide'}
rugged_model <- stan(model_code = stan_program,
                      data = stan_data,
                      warmup = 100,
                      chains = 4,
                      control = list(adapt_delta = 0.8))
```

```{r, message=FALSE, results='hide'}
rugged_model2 <- stan(model_code = stan_program,
                      data = stan_data,
                      warmup = 250,
                      chains = 4,
                      control = list(adapt_delta = 0.8))
```

```{r, message=FALSE, results='hide'}
rugged_model3 <- stan(model_code = stan_program,
                      data = stan_data,
                      warmup = 500,
                      chains = 4,
                      control = list(adapt_delta = 0.8))
```

```{r, message=FALSE, results='hide'}
rugged_model4 <- stan(model_code = stan_program,
                      data = stan_data,
                      warmup = 750,
                      chains = 4,
                      control = list(adapt_delta = 0.8))

```

```{r}
rugged_model
rugged_model2
rugged_model3
rugged_model4
```

Rugged_model - warmup = 100
sigma n_eff = 2940
a n_eff = 2291
b n_eff = 2291
c n_eff = 1281
d n_eff = 1193
lp__ n_eff = 1291

Rugged_model2 - warmup = 250
sigma n_eff = 2729
a n_eff = 2703
b n_eff = 2981
c n_eff = 2430
d n_eff = 2301
lp__ n_eff = 1266

Rugged_model3 - warmup = 500
sigma n_eff = 1583
a n_eff = 2282
b n_eff = 2168
c n_eff = 1610
d n_eff = 1343
lp__ n_eff = 846

Rugged_model4 - warmup = 750
sigma n_eff = 723
a n_eff = 1055
b n_eff = 827
c n_eff = 748
d n_eff = 757
lp__ n_eff = 465

It appears that 250 warmup got the most n_eff and also had rhat = 1 so about 25% of the total samples as warmup seems to be enough.