---
title: "Stat 344 -- PS 2"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too
library(tidyverse)
library(mosaic)
library(maxLik)
library(ggplot2)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->


```{r}
mrc <- read.csv('http://sldr.netlify.app/data/mrc.csv')
```


### Problem 5.6

X ~ Multinom($\pi$)

$n = \sum x_i$

$l(\pi ; x) = logL(\pi ; x) = \sum _i log(dmultinnom(x_i, \pi)$)

$\infty x_1log(\pi_1) + ... + x_klog(\pi_k) = \tilde{l}(\pi; x)$
half infinity, not full infinity

Given $g(\pi) = \sum_{i=1}^k \pi_i$ when $g(\pi) =1$

$\Delta \tilde{l} (x; \pi) = \lambda \Delta g(\pi)$

$g(\pi) = 1$

$\frac{x_1}{\pi_1} = \lambda_1$ -> $\pi_1 = \frac{x_1}{\lambda}$

$\frac{x_2}{\pi_2} = \lambda_1$ -> $\pi_2 = \frac{x_2}{\lambda}$

$\frac{x_1}{\pi_3} = \lambda_3$ -> $\pi_3 = \frac{x_3}{\lambda}$

Substituting in, $\pi_1 + \pi_2 + \pi_3 +...+ \pi_k = 1$

$\frac{x_1 + x_2 + x_3 + ... + x_k}{\lambda} = 1$

$\frac{n}{\lambda} = 1$ -> $\lambda = n$

So then $\pi_1 = \frac{x_1}{n}, \pi_2 = \frac{x_2}{n}, ..., \pi_k = \frac{x_k}{n}$ so $\pi_i = \frac{x_i}{n}$


### Problem 5.10

Repeat Example 5.1.4 Using numerical methods

Find the Maximum likelihood estimator for 


```{r, echo = TRUE}
#$L(\theta ; x) = 0$ 
  
#if $\theta < 8.7$ , 

#$(\frac{1}{\theta})^n $ 

#if $ \theta \ge 8.7$  
```



$log(L(\theta ;x)) = log(\Pi_{i=1} ^6[(\frac{1}{\theta})^n])$

$=log(\frac{1}{\theta})^6$

$=\frac{d}{d\theta} (6log(\frac{1}{\theta})) = -\frac{6}{\theta}$ 

$-\frac{6}{\theta} = 0$

$\hat{\theta} = \frac{1}{6}$

```{r}
data <- c(1.6, 2.8, 6.2, 8.2, 8.5, 8.7)
```


```{r}
LL <- function(theta, x){
  b <- theta[1]
  dunif(x, min = 0, max = b, log = TRUE)
}
```

```{r}
maxLik(LL, start = c(b = 10), x = data)
```

### Problem 5.12

```{r}
mrc <- read.csv('http://sldr.netlify.app/data/mrc.csv') |>
  filter(mobility > 0)
```

```{r}
gf_histogram(~mobility, data = mrc)
```

```{r}
LL2 <- function(theta, x){
  alpha <- theta[1]
  beta <- theta[2]
  if (alpha < 0) return(NA)
  if (beta < 0) return(NA)
  dbeta(x, shape1 = alpha, shape2 = beta, log = TRUE)
}
```

```{r}
maxLik(LL2, start = c(alpha = 1, beta = 150), x = mrc$mobility)
```

```{r}
gf_dist("beta", shape1 = 3.155, shape2= 192.61)
```

The model seems pretty good for the model given how well the beta distribution with the given parameters from the MLE function. When overlayed a histogram of the original data, the graphs appear to line up pretty well in distribution. 




  
