---
title: "Stat 344 -- PS 03"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too
library(tidyverse)
library(mosaic)
library(maxLik)
library(ggplot2)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Problem 5.24

a.
See attached pdf.

b. 
```{r}
x <- c(.64, .92, .73, .96, .98, .33, .8, .96, .81, .76, .98, .75, .87, .82, .44, .96, .61, .32, .67, .98, .96, .88, .85, 1, .86, .88, .80, .83, .64, .5)
```


```{r}
W <- function(x, theta0 = 0){
  x_sum <- sum(x) 
  n <- length(x) 
  theta.hat <- (n / -sum(log(x)) - 1)
  -2 * log(theta0 + 1) + 2 * log(theta.hat + 1) + (log(x_sum) * (2 * theta.hat - 2 * theta0))
}
```

```{r}
W(x)
```
c. 
```{r}
p_value <- 1 - pchisq(W(x), df = 1)
p_value
```

There doesn't seem to be any reason to be concerned about this p-value since it is so small to reject the null hypothesis that theta = 0 when the data doesn't seem to suggest that.

d.
```{r}
rdata <- do(5000) * runif(30, 0,1)

statTally(x, rdata, W)
```
Simulated data reveals a p-value of p = 0.000287 which is consistent with the real data.

e.
The two p-values are slightly different but both are really small and consistent with each other. Both reject the null hypothesis that theta = 0 because the data shows that theta isn't really even close to 0.

### Problem 5.29

See attached hand-written pdf

### Problem 5.30

a.

theta = p(1)^n1 * p(2)^n2 * p(3)^n3 * p(4)^n4
      = .25^4(2theta - 3theta^2 + theta^4)^n
```{r}
n <- c(1997, 906, 904, 32)

LL <- function(theta, x){
  x[1] * log(.25 * (2 + theta)) +  x[2] * log(.25 * (1-theta)) + x[3] * log(.25 * (1-theta)) + x[4] * log(.25 * (theta))
}

maxLik(LL, start = .5 , x = n)
```
      
MLE = 0.0357123 

b.

```{r}
W2 <- 2 * (LL(0.0357123, n) - LL(0.05, n))

p_value2 <- 1 - pchisq(W2, df = 1)
p_value2
```
P-value = 0.03 therefore reject the hypothesis that theta = 0.05

c.

```{r}
W3 <- 2 * (LL(0.0357123, n) - LL(0.03, n))

p_value3 <- 1 - pchisq(W3, df = 1)
p_value3
```
P-value = 0.318 therefore fail to reject the hypothesis that theta = 0.03

d.

```{r}
W4 <- 2 * (LL(0.0357123, n) - LL(0.07, n))

p_value4 <- 1 - pchisq(W4, df = 1)
p_value4
```

P-value = 0.0000039 therefore reject the hypothesis that theta = 0.07
