---
title: "Stat 344 -- PS05"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too
library(tidyverse)
library(mosaic)
library(maxLik)
library(ggplot2)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Problem 5.26



### 5.30

```{r}
n <- c(1997, 906, 904, 32)

LL <- function(theta, x){
  x[1] * log(.25 * (2 + theta)) +  x[2] * log(.25 * (1-theta)) + x[3] * log(.25 * (1-theta)) + x[4] * log(.25 * (theta))
}

maxLik(LL, start = .5 , x = n)
```

e.

```{r}
o <- n; o

sn <- sum(n); sn

theta <- 0.0357123

probs <- c(.25*(2 + theta), (.25 * (1-theta)), (.25 * (1-theta)), (.25 * theta))

e <- probs * sn

G <- 2 * sum(o * log(o / e)); G

1 - pchisq(G, df = 2)
```
Data did not fit the model well with a goodness-of-fit p-value of .36

f.

```{#r}

theta.hat <- 0.0357123 

pval_minus_critical <- function(theta0) {
  2 * (LL(theta.hat, x) - LL(theta0, x)) - qchisq(.95, df = 1)
}

lo <- uniroot(pval_minus_critical, c(0, theta.hat)) |> value()
hi <- uniroot(pval_minus_critical, c(theta.hat, 1)) |> value()

c(lo,hi)
```


### 5.33

```{r}
bin.pois <- function() {
  
rdata <- data.frame(V1 = rpois(1000, lambda = 15))

breaks <-c(0, 5, 10, 15, 20, 25, Inf)

rdata <- rdata |> 
  mutate(binned = cut(V1, breaks = breaks))
tally(~binned, data = rdata)

count_dat <- data.frame(tally(~binned, data = rdata))
count_dat <- count_dat |>
mutate(probs = diff(ppois(breaks, lambda = 10)),
e = sum(count_dat$Freq) * probs )

G_pois <- 2 * sum( count_dat$Freq * log( count_dat$Freq / count_dat$e))

1 - pchisq(G_pois, df = nrow(count_dat) - 1 - 2)
}

bin.pois()
```
  
### 5.36

```{r}
Fusion1m <- merge(FUSION1, Pheno, by = "id", all = FALSE)

genotype <- tally(~ t2d + genotype, data = Fusion1m)

Gdose <- tally(~ t2d + Gdose, data = Fusion1m)
```

a. T is the most common allele among cases than among controls.

b. 

```{r}
chisq.test(genotype)
```

Based on the chi-squared test, there is an association between the SNP and type 2 diabetes.

c. knowing this marker was just one part of a larger study with over 300,000 SNPs changes how to interpret these results because
