---
title: "Stat 344 -- PS06"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
    latex_engine: xelatex
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
always_allow_html: yes
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Problem 5.46

a.
P(X|λ) = (e^(-λ) * λ^Σxᵢ) / (x₁! * x₂! * ... * xₙ!)

gamma pdf

f(λ|α, θ) = (1 / (θ^α * Γ(α))) * λ^(α-1) * e^(-λ/θ)

Apply Bayes

P(λ|X) ∝ P(X|λ) * P(λ)

Log both sides

log(P(λ|X)) ∝ log(P(X|λ)) + log(P(λ))

Then substitute the right side of the equation

log(P(λ|X)) ∝ Σxᵢ * log(λ) - λ - log(x₁! * x₂! * ... * xₙ!) + (α - 1) * log(λ) - λ/θ - log(θ^α * Γ(α))

Thus,

P(λ|X) ~ gamma(α + Σxᵢ, n + 1/θ)

b.

E(λ) = α * β, so given  P(λ|X) ~ gamma(α + Σxᵢ, n + 1/θ),

E(λ|X) = (α + Σxᵢ) * (n + 1/θ)

c.

As n approaches infinity, so does the posterior distribution

d.

```{r}
classes <- c(9, 39, 22, 35, 28)

NormalGrid <-
  expand.grid(
    lambda = seq(0, 50, length.out = 100000)
  ) |>
  mutate(
    prior = dgamma(lambda, shape = 3, rate = 1/3), ## prior choices
    likelihood = mapply(function(l) {prod(dpois(classes, lambda = l))}, l = lambda),
    posterior = prior * likelihood /  sum(prior * likelihood) * 1000
  )

NormalGrid |> arrange(-posterior) |> head(3)

PosteriorSample <-
  sample(NormalGrid, size = 1e5, replace = TRUE, prob = NormalGrid$posterior)

gf_histogram(~lambda, data = PosteriorSample, binwidth = 0.2)

gf_line(prior ~ lambda, color = ~ "prior", data = NormalGrid) |>
  gf_line(posterior ~ lambda, color = ~ "posterior", data = NormalGrid)

cdata(~ lambda, data = PosteriorSample, p = 0.95)

```



### Problem 5.47

```{r}
Positions <- rgeo(25)
leaflet_map(position = Positions, mark = TRUE)
water <- 16
land <- 9
```

```{r}
qbeta(c(0.05, 0.95), 17 , 10 )
```
0.4738376 0.7743001 90% credible interval with a uniform prior



### Problem 5.49

P(X|tau) = (1 / (sqrt(2π/tau))^n) * e^(-Σ(xᵢ^2) * tau/2)

pdf: f(tau|alpha, beta) = (1 / (beta^alpha * Γ(alpha))) * tau^(alpha-1) * e^(-tau * beta)

Bayes: P(tau|X) ∝ P(X|tau) * P(tau)

Log: log(P(tau|X)) ∝ log(P(X|tau)) + log(P(tau))

Substitute: log(P(tau|X)) ∝ -Σ(xᵢ^2) * tau/2 + (alpha - 1) * log(tau) - tau * beta - log(beta^alpha * Γ(alpha))

log(P(tau|X)) ∝ -(Σ(xᵢ^2)/2 - beta) * tau + (alpha - 1) * log(tau) - log(beta^alpha * Γ(alpha))

Thus, P(tau|X) ~ gamma(alpha + n/2, (Σ(xᵢ^2)/2 - beta))

  
