---
title: "Stat 344 -- RMarkdown Challenge"
author: "Luca Seazzu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Mathematical Notation   

The probability of both event $A$ and event $B$ happening is $P(A \cap B)$, and if $A$ and $B$ are independent, this quantity is equivalent to $P(A)P(B)$.
One way of writing the equation for a simple linear regression of $y$ as a function of $x$ is:

$$y = \beta_0 + \beta_x + \epsilon,$$

$$\epsilon \sim N(0, \sigma)$$

### A Dataset

A dataset on participants in the Olympic Games is available at https://sldr.netlify.com/data/athlete_events.csv

A glimpse() of the dataset is below:

```{r, include = FALSE}
olympics <- read.csv('https://sldr.netlify.com/data/athlete_events.csv')
```


```{r}
glimpse(olympics)
```

### Graphics

#Age distribution of Olympians
```{r}
gf_dhistogram(data = olympics, ~Age)
```

#Numbers of Male and Female Olympians
```{r}
gf_props(~Sex, data = olympics)
```

#Extra Credit: Proportion Female by Year
```{r}
gf_point(prop(~Sex == 'F') ~Year, data = olympics)
```
Couldn't quite remember how to do this off the top of my head.

#Extra Credit: Total US Medals, 1896-2016
```{r}

```

Also couldn't quite remember how to do this but it is something like tally I think.



  
